Aa Language Reference Manual

### 2023-09-16 v2/Ahir-asm/devel/docs/AaLRM.tex
Small fix:    when  a guarded statement is present in AA      $guard (op) p := (q + r)    the latency of the operation (q+r) at the point    of the guard increases by 1.  This needs to    be accounted for in pipeline balancing.
   Further, the rate behaviour of this statement
   in a pipeline will be 1/2 capacity unless the
   statement is double buffered..

   This has been fixed by making the guarded (in this case, add)
   operator have a double buffered output wire.  The fix
   is applied in printing the vC.

The rules for scoping are as follows:
\begin{enumerate}
\item A declared variable defined in a scope is visible
in all descendent scopes. 
\item A reference to a variable $b$ in a scope X is resolved by checking
whether the variable is defined in that scope, and if not found there,
by checking in the scope that contains the scope X, and so on.
\item A scope can read from variables that are defined in descendant
scopes.  
\item A scope can export a variable defined within its scope to
its parent scope.   The name  of the exported variable must not
clash with the name of any other variable visible in the
parent scope.
\item A scope can read from variables that are defined in ancestor scopes.
\item A scope can only write to one of the following:
\begin{itemize}
\item an implicit variable defined in the scope.
\item a storage or pipe variable defined in the scope or an ancestor
of the scope.
\item an output argument of a module of which the scope is a descendant.
In this case, there can be at most one statement in the entire module
which writes to this output argument.
\end{itemize}
\end{enumerate}

A variable reference in a statement may be specified 
as follows
\begin{verbatim}
a            look for variable a in current scope; 
             if not found look in the parent.

:a           same as the previous case

../:a        look for variable a starting from the parent of
             the current scope.

%p:a         look for variable a starting from the child scope 
             with label p (child scope of the current scope).

../../:a     look for variable starting from the parent of
             the parent of the current scope.

%p%q:a       look for variable starting from the child scope 
             with label q of the child scope with label p of 
             the current scope

a[10]        look for a storage variable a defined in the 
             current scope.  If not found, look for it in
             the parent scope, and so on.  If eventually found, 
             access the corresponding element of the composite
             object.
\end{verbatim}
Thus, a generic variable reference has the form 
\begin{verbatim} 
    scope-reference : variable-reference 
\end{verbatim} 
and the scoping rules forbid a scope from accessing variables
which are not defined in either an ancestor or a descendant
of the scope (the export mechanism should be used if this
is needed).

@@

>>>>

The rules for scoping are as follows:
\begin{enumerate}
\item A declared variable defined in a scope is visible
in all descendent scopes. 
\item A reference to a variable $b$ in a scope X is resolved by checking
whether the variable is defined in that scope, and if not found there,
by checking in the scope that contains the scope X, and so on.
\item A scope can read from variables that are defined in ancestor scopes.
\item A scope can only write to one of the following:
\begin{itemize}
\item an implicit variable defined in the scope.
\item a storage or pipe variable defined in the scope or an ancestor
of the scope.
\item an output argument of a module of which the scope is a descendant.
In this case, there can be at most one statement in the entire module
which writes to this output argument.
\end{itemize}
\end{enumerate}

If, from within a scope, we wish to read the value of
a variable defined in a child scope, then the child scope
must export the value of the variable so that the parent
scope can access it.

@@

\section{Gated clocks}

	Gated clocks can be declared and controlled within the Aa
	language framework.   There are two control scenarios which
	are supported.
\begin{itemize}
\item  Automatic mode:
	   the gated clock can be attached to a module, and is
	  enabled whenever the module is activated (by the
	  start\_req/start\_ack pair), and disabled when the
	  module is no longer active, i.e., it has no further
	  jobs in progress inside it.
\item Explicit mode:
	   the gated clock enable signal is generated from within
	   some logic described in Aa.
\end{itemize}

\subsection{Creating and using an explicitly controlled gated clock}

	To create an explicit gated clock, we need to declare an enable signal
	and a gated clock to which it is matched.  For example
\begin{verbatim}
// signal declaration of an enable.
$pipe  data_1_clock_enable: $uint<1> $signal
// gated clock attached to enable.
$gated_clock  data_1_clock data_1_clock_enable
\end{verbatim}

	Now the declared gated clock can be applied to
	a module foo by the statement
\begin{verbatim}
	$use_gated_clock foo  data_1_clock
\end{verbatim}
 	It is the responsibility of the programmer
	to control the enable signal (data\_1\_clock\_enable above).
	This enable signal will control the gated clock data\_1\_clock
	which is used by the module foo.

\subsection{Declaring and using automatic gated clocks}

	In this case, all we need to declare is that 
	a module ``bar'' will run on a gated clock which
	is enabled when the module is working on an active job
	and disabled otherwise.  This is done by the
	statement
\begin{verbatim}
$use_gated_clock bar
\end{verbatim}
